---
title: 'Part 1: Simulation Exercise Instructions'
output:
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference Course Project 1


## Overview
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. Note that I will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. I should

Show the sample mean and compare it to the theoretical mean of the distribution.
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
Show that the distribution is approximately normal.

---
## Simulations

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
#load libraries
library(ggplot2)

#set constants
lambda <- 0.2 # lambda for rexp
n <- 40 # number of exponetials
number_of_tests <- 1000 # number of tests

#set the seed to create reproducability
set.seed(42)

#run the test resulting in n x number_of_tests matrix
exponential_distributions <- matrix(data=rexp(n * number_of_tests, lambda), nrow=number_of_tests)
exponential_distributions_mean <- data.frame(means=apply(exponential_distributions, 1, mean))

# plot the means
ggplot(data = exponential_distributions_mean, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(exponential_distributions_mean$means), max(exponential_distributions_mean$means), by=1)))


```


## Sample Mean Vs Theoretical Mean

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
sample_mean <- 1/lambda
theoretical_mean <- mean(exponential_distributions_mean$means)

```

* Sample Mean : `r sample_mean`
* Theoretical Mean : `r theoretical_mean`


## Sample Variance versus Theoretical Variance

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
sample_sd <- sample_mean / sqrt(n)
sample_variance <- sample_sd^2

theoretical_sd <- sd(exponential_distributions_mean$means)
theoretical_variance <- var(exponential_distributions_mean$means)

```
* Sample deviation : `r sample_sd`
* Theoretical deviation : `r theoretical_sd`
* Sample Variance : `r sample_variance`
* Theoretical Variance : `r theoretical_variance`


## Distribution
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
#plot the means
ggplot(data = exponential_distributions_mean, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = sample_mean , sd = sample_sd), colour = "red", size=1) + 
  geom_vline(xintercept = sample_mean, size=1, colour="#CC0000") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = sample_variance, size=1, colour="#0000CC") + 
  scale_x_continuous(breaks=seq(sample_mean-3,sample_mean+3,1), limits=c(sample_mean-3,sample_mean+3)) 
```
